---
title: "04. Stylophora cross-stages comparison"
date: "`r format(Sys.time(), '%d %B, %Y')`"
auhor: "Anamaria Elek"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
    toc_depth: 2
---
```{r options, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, fig.align = 'center')
```

# Set up the analysis

Load required packages and functions.

```{r setup, eval=TRUE, message=FALSE, warning=FALSE}
library(data.table)
library(stringr)
library(ape)
library(tidytree)
library(ggtree)
library(treeio)
library(dendextend)
library(phangorn)
library(phytools)
library(ggplot2)
theme_py <- theme_classic() + theme(
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_rect(colour = "black", fill = NA),
  text = element_text(size=20)
)
theme_set(theme_py)
# load functions for downstream analysis
scripts <- list.files("metacell_downstream_functions",full.names=TRUE)
for (file in scripts) source(file)
```

# Cell type tree

```{r}
outdir <- "cross_stages/tree"
dir.create(outdir, showWarnings=FALSE)
```

Load previously generated metacell expression data summarized at cell type level.

```{r ct}
adult_mcfp <- read.table("clustering_coral/scdb/Spis_coral_cell_type_gene_FC")
colnames(adult_mcfp) <- str_replace(colnames(adult_mcfp),"\\.","-")
adult_annot_file <-"clustering_coral/scdb/Spis_coral_cell_type_annotation"
adult_annot <- fread(adult_annot_file)
adult_mcfp <- adult_mcfp[,adult_annot$cell_type]

larva_mcfp <- read.table("clustering_larva/scdb/Spis_larva_cell_type_gene_FC")
larva_annot_file <-"clustering_larva/scdb/Spis_larva_cell_type_annotation"
larva_annot <- fread(larva_annot_file)
larva_mcfp <- larva_mcfp[,larva_annot$cell_type]

polyp_mcfp <- read.table("clustering_polyp/scdb/Spis_polyp_cell_type_gene_FC")
polyp_annot_file <-"clustering_polyp/scdb/Spis_polyp_cell_type_annotation"
polyp_annot <- fread(polyp_annot_file)
polyp_mcfp <- polyp_mcfp[,polyp_annot$cell_type]
```

Combine cell type annotations.

```{r ct_anns, eval=TRUE}
ct_annot1=read.table("clustering_coral/scdb/Spis_coral_cell_type_annotation",header=TRUE,stringsAsFactors=FALSE)
ct_annot2=read.table("clustering_polyp/scdb/Spis_polyp_cell_type_annotation",header=TRUE,stringsAsFactors=FALSE)
ct_annot3=read.table("clustering_larva/scdb/Spis_larva_cell_type_annotation",header=TRUE,stringsAsFactors=FALSE)
ct_ann_dt <- rbindlist(list(Adult=ct_annot1,Polyp=ct_annot2,Larva=ct_annot3),idcol="sp")
ct_ann_dt[,metacell:=paste(sp,metacell,sep="_")]
ct_ann_dt[,cell_type:=paste(sp,cell_type,sep="_")]
ct_ann_dt[,c("sp"):=NULL]
ctcol=c(ct_annot1$color,ct_annot2$color,ct_annot3$color)
names(ctcol)=c(
  paste0("Adult_",ct_annot1$cell_type),
  paste0("Polyp_",ct_annot2$cell_type),
  paste0("Larva_",ct_annot3$cell_type)
)
fwrite(ct_ann_dt,file.path(outdir,"Spis_all_stages_cell_type_annotation"),sep="\t")
```

Combine expression data for stages. To avoid loosing stage-specific variable genes, we assign them value 1 in datasets in which they are missing.

```{r combine_all, eval=FALSE}
all_genes <- unique(c(rownames(adult_mcfp), rownames(polyp_mcfp), rownames(larva_mcfp))) # 31398
add_missing <- function(mcfp,all_genes) {
  missing <- setdiff(all_genes,rownames(mcfp))
  addition <- matrix(1,nrow=length(missing),ncol=ncol(mcfp))
  rownames(addition) <- missing
  colnames(addition) <- colnames(mcfp)
  rbind(mcfp,addition)[all_genes,]
}
adult_mat <- add_missing(adult_mcfp,all_genes)
colnames(adult_mat) <- paste0("Adult_", colnames(adult_mat))
polyp_mat <- add_missing(polyp_mcfp,all_genes)
colnames(polyp_mat) <- paste0("Polyp_",colnames(polyp_mat))
larva_mat <- add_missing(larva_mcfp,all_genes)
colnames(larva_mat) <- paste0("Larva_",colnames(larva_mat))
mc_fp <- as.matrix(cbind(cbind(adult_mat, polyp_mat), larva_mat))
write.table(mc_fp,file.path(outdir,"Spis_all_stages_gene_FC.tsv"))
```

Define variable genes.

```{r var_genes, eval=TRUE}
mc_fp <- read.table(file.path(outdir,"Spis_all_stages_gene_FC.tsv"))
colnames(mc_fp) <- str_replace(colnames(mc_fp),"\\.","-")
var_genes_fc <- 1.8 
var_genes <- names(which(apply(mc_fp,1,function(x) sort(x,decreasing=T)[2])>var_genes_fc)) # 5881
```

There are `r length(var_genes)` variable genes in the three datasets. We use these genes for hierarchical clustering and building the cell type tree.  
Now we build cell type trees and generate the final tree from the co-occurence matrix.

```{r cooccur_tree, eval=FALSE}
# Co-occurence tree
set.seed(4343)
h <- c(0.75,0.95)
p <- 0.75 
clustering_algorithm <- "hclust"
clustering_method <- "average"
cor_method <- "pearson"
treename <- sprintf(
  "cross_stages_ct_vargenes%s_cooc%sp%s%sh%s",
  var_genes_fc,clustering_method,p,clustering_algorithm,paste(h,collapse=",")
)
tree_vargenes_cooc <- treeFromEnsembleClustering(
  x=mc_fp, vargenes=var_genes, p=p, h=h,  bootstrap=FALSE,
  clustering_algorithm=clustering_algorithm, clustering_method=clustering_method, 
  cor_method=cor_method
)
tree_rds <- file.path(outdir,sprintf("%s.RDS",treename))
tree <- tree_vargenes_cooc$tree
saveRDS(tree,tree_rds)
saveTreeFiles(
  tree_rds, ctcol=ctcol, out_dir = outdir, save_nexus = TRUE, scale_branches = FALSE,
  width_tree = 10, height_tree = 24, width_polytree = 10, height_polytree = 22
)

# Co-occurrence matrix
col=colorRamp2(
  c(0,100,200,400,600,700,800,900,1000),
  colors=c(c("white",'#ffffe5','#fff7bc','#fee391','#fec44f','#fe9929','#ec7014','#cc4c02','#990000'))
)
plotTreeMatrix(
  mat = tree_vargenes_cooc$cooccurrence, tree = tree, ctcol=ctcol, hmapcol = col,
  name="co-occurence", out_name = file.path(outdir, sprintf("matrix_%s.pdf",treename)),
  width = 20, height = 20
)
      
```

We reorder the tree nodes externally (FigTree), load it back to R to save nexus file with colors and RDS file.

```{r eval=FALSE, fig.height=18}
tree_file=file.path(outdir,"cell_type_tree_vargenes1.8_coocaveragep0.75hclusth0.75,0.95_politomies0.10.beast")
tree <- as.phylo(read.beast(tree_file)); plot(tree,tip.color=ctcol[tree$tip.label])
treeColorTips(
  tree_file=tree_file, ctcol=ctcol,
  out_file=file.path(outdir,"cell_type_tree_vargenes1.8_coocaveragep0.75hclusth0.75,0.95_politomies0.10.beast")
)
saveRDS(tree,str_replace(tree_file,"beast","RDS"))
```

Also reorder co-occurance matrix by final tip ordering.

```{r}
plotTreeMatrix(
  mat = tree_vargenes_cooc$cooccurrence, tree = tree, build_tree=FALSE, rev=FALSE,
  ctcol=ctcol, hmapcol = col, name="co-occurence", 
  out_name = str_replace_all(tree_file,c("cell_type"="matrix_cell_type","beast"="pdf")),
  width = 12, height = 12, row_names_side = "left"
)
```

# Gap genes

Accessing internal tree nodes to find feature (gap) genes. This is `scripts/gapgenes.R` script, submitted to HPC using `gapgenes.sh` submission script.

The reults can be interactively explored in the [shiny app](https://sebe-lab.shinyapps.io/Stylophora_cell_type_tree).

# GO enrichment
We compare calicoblast genes in adult and polyp, as well as gastrodermis and symbiodinium hosting cells, using GO enrichment.

```{r warning=FALSE, message=FALSE}
source("scripts/topGO.R")
gomap <- readRDS("annotation/gomap_spis.RDS")
tgalg <- "classic"
godir <- file.path("cross_stages","go",tgalg); dir.create(godir,recursive=TRUE,showWarnings=FALSE)

# gene lists of genes upregulated in calicoblasts, and in gastrodermis with and without symbiodinum
gls_files <- list.files("cross_stages/genelists/",pattern=".*txt",full.names=TRUE)
names(gls_files) <- str_remove(basename(gls_files),"\\.txt")
cts <- names(gls_files)

# calculate GO enrichment
godt <- rbindlist(sapply(names(gls_files), function(ct) {
  list_interest <- fread(gls_files[ct],header=FALSE)[[1]]
  setDT(topgofun(
    list_interest, gomap, tg_algorithm=tgalg, output_name=file.path(godir,sprintf("go.%s",ct)),
    name_geneset=ct, ontology_set=c("MF","CC","BP"), firstSigNodes=20
  ))
}, USE.NAMES=TRUE, simplify=FALSE), idcol="group")
fwrite(godt,file.path(godir,sprintf("topGO.%s.tsv",tgalg)),sep="\t")

# select significant results
sign_go <- godt[,.SD[order(pval_adj,decreasing=FALSE)][pval_adj<0.05],group]
sign_go[,xid:=sprintf("%s:%s",GO.ID,Term)]
sign_go[,logpadj:=-1*log10(pval_adj)]

for (ct in cts) {
  godt <- sign_go[group==ct]
  setorder(godt,"Significant","pval_adj")
  godt[,xid:=factor(xid,levels=unique(xid))]
  if (nrow(godt)>0) {
    gp <- ggplot(godt, aes(xid, Significant, fill=pval_adj)) +
      geom_bar(stat="identity") + coord_flip() + 
      facet_grid(ontology~.,scales="free_y",space="free_y",drop=TRUE) +
      scale_y_continuous(expand = c(0,0)) +
      scale_fill_gradient(limits = c(0,0.05), low=grey.colors(3)[2], high=grey.colors(3)[1], breaks=c(0,0.05)) +
      labs(x="", y="significant genes", fill="adjusted\np value") + 
      theme(strip.background=element_blank())
    # awful helper function to try estimating plot dimensions
    .plotdims <- function(x) {
      if (x<6) {c(3,10)} else if (x<11) {c(10, 16)} else if (x<21) {c(12,15)} else if (x<41) {c(14,14)} else if (x<51) {c(16,13)} else if (x<100) {c(40,14)} else if (x<400) {c(60,15)} else if (x<600) {c(80,14)} else {c(100,14)}
    }
    dims <- .plotdims(nrow(godt))
    ph = dims[1]; pw = dims[2]
    pdf(file.path(godir,sprintf("go%s_plot_%s.pdf",tgalg,ct)),height=ph,width=pw,useDingbats=TRUE)
    print(gp)
    dev.off()
  }
}
```

