---
title: "07. Hydra vulgaris metacell analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
auhor: "Anamaria Elek"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
    toc_depth: 2
---
```{r options, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, fig.align = 'center')
```

# Set up metacell analysis

Load required packages and functions.

```{r setup, eval=TRUE, message=FALSE, warning=FALSE}
library(data.table)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(ggrastr)
library(plotly)
library(scales)
library(stringr)
library(RColorBrewer)
library(circlize)
library(tgconfig)
library(metacell)
theme_py <- theme_classic() + theme(
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_rect(colour = "black", fill = NA),
  text = element_text(size=20)
)
theme_set(theme_py)
# load functions for downstream analysis
scripts <- list.files("metacell_downstream_functions",full.names=TRUE)
for (file in scripts) source(file)
```

Initialize database and figures output directory.

```{r mc_setup, eval=TRUE, warning=FALSE, message=FALSE}
require(metacell)

# analysis dir
cluster_dir <- "clustering_hydra"
dir.create(cluster_dir, showWarnings=FALSE)

# initialize db directory
scdb_dir <- file.path(cluster_dir, "scdb")
dir.create(scdb_dir, showWarnings=FALSE)
scdb_init(scdb_dir, force_reinit=TRUE)

# plotting directory
scdb_fig_dir <- file.path(scdb_dir,"figs")
dir.create(scdb_fig_dir, showWarnings=FALSE)
scfigs_init(scdb_fig_dir)

# directory with annotations
annot_dir <- "annotation_hydra"
```

Download UMI counts matrix from Siebert et al. (Science 2019) and 

```{r siebert_umi}
require(Seurat)
require(data.table)
require(stringr)
seurat <- readRDS(file.path(cluster_dir,"Hydra_Seurat_Whole_Genome.rds"))
df <- as.data.table(as.matrix(seurat@raw.data), keep.rownames = "")
setnames(df, "", "name")
# get transcript id
df[,id:=sprintf("%s_1", str_extract(name,pattern="g\\d+"))][
  # get transcript number
  ,num:=str_extract(name,pattern="t\\d")][
    # get blast info
    ,blast:=str_extract(name,pattern="(?<=\\|).+")][
      # total number of transcripts for a gene
      ,numtot:=.N,id]
# number of genes with more than one transcript
dups <- unique(df[numtot>1][,id])
length(dups)
# UMI counts per gene
df <- df[,umitot:=rowSums(.SD), .SDcols=2:length(seurat@cell.names)][
  ,.SD[which.max(umitot)], by=id]
# get just UMI matrix now
umi_mat <- df[,-c("name","num","blast","numtot","umitot")]
setnames(umi_mat, "id", "")
# save UMI matrix
umi_table_file <- file.path(annot_dir, "Hydra_Seurat_Whole_Genome_filt_dup.tsv")
fwrite(umi_mat, umi_table_file, sep = "\t")
```

Also get association of cells to cell types from  Siebert et al. (Science 2019).

```{r siebert_ct}
dt <- data.table(
  cell=names(seurat@ident),
  cell_type=seurat@ident
)
cell_annot_file <- file.path(annot_dir, "Hydra_Seurat_Whole_Genome_cell_annot.tsv")
fwrite(dt, cell_annot_file, sep="\t")

dt <- fread(cell_annot_file, sep="\t")
mc_dt <- fread(file.path(scdb_dir, "hydra_mc_assigned_ct.csv"))[,.(cell,mc)]
scr_transfer_annotations(dt=dt, mc_dt=mc_dt)
```

# Metacell piepline 

Load the expression data to metacell database.  

```{r mcell_import}
# import UMI table
mat_nm <- "hydra"
mcell_import_scmat_tsv(
  mat_nm = mat_nm,
  fn = umi_table_file,
  dset_nm = "siebert2019")
mat <- scdb_mat(mat_nm)
```

Filter the expression data and calculate gene statistics.

```{r mcell_filter}
# filtering by cell depth (UMIs/cell)
# already filtered for cells with 300 to 7000 detected genes and 500 to 50000 UMIs
x=colSums(as.matrix(mat@mat))
range(x)
mcell_plot_umis_per_cell('hydra', min_umis_cutoff = 500)

# filtering ERCCs 
# no ERCCs in this dataset, already filtered
erccs=rownames(mat@mat)[grepl("ERCC",rownames(mat@mat))] 
mcell_mat_ignore_genes(new_mat_id="hydra",mat_id="hydra",ig_genes=erccs)

# genes' statistics
mcell_add_gene_stat(gstat_id="gstat", mat_id="hydra", force=T)

# reload matrix
mat=scdb_mat("hydra")
```

Select marker genes for building metacell graph.  

```{r mcell_markers}
mcell_gset_filter_multi(gstat_id="gstat",gset_id="clust_markers",T_tot=30,T_top3=2,T_szcor=-0.08,T_niche=0.05,force_new=T)
mcell_plot_gstats(gstat_id="gstat", gset_id="clust_markers")
```

Building metacells.  

```{r mcell_knn}
# build cell graph
mcell_add_cgraph_from_mat_bknn(mat_id="hydra",gset_id="clust_markers",graph_id="graphk100",K=100,dsamp=F)

# resample graph and compute metacells for each resampling
mcell_coclust_from_graph_resamp(coc_id="coc500_min20",graph_id="graphk100",min_mc_size=15,p_resamp=0.75,n_resamp=500)

# final graph and metacells from co-occurence in resampled graphs
mcell_mc_from_coclust_balanced(mc_id="mc",coc_id="coc500_min20",mat_id="hydra",K=100,min_mc_size=15,alpha=2)
```

# Transfer annotations 

```{r transfer_annot_files}
mc_cell_info_file <- file.path(scdb_dir, "hydra_cell_annotation_transfered.tsv")
mc_annotation_file <- file.path(scdb_dir, "hydra_metacell_annotation.tsv")
```

Transfer annotations from Siebert et al. (Science 2019).

```{r transfer_annot}
mc_dt <- data.table(
  cell=names(mc@mc),
  mc=mc@mc
)
dt <- fread(cell_annot_file) # defined earlier
dt[mc_dt, on="cell", mc:=i.mc] # add mc annotation
dt <- dt[!is.na(mc)] # remove cells that are not assigned to any mc
dt[,cells_from_mc_in_ct:=.N,by=.(mc,ct)] # for each mc, count how many cells in it are assigned to each ct
dt[,max_cells_ct:=max(cells_from_mc_in_ct),mc] # find most occuring ct for each mc
dt[,max:={cells_from_mc_in_ct==max_cells_ct}]
max_ct <- unique(dt[max==TRUE,.(ct,mc)])
all.equal(dim(max_ct)[1], dim(mc@mc_fp)[2]) # check for 50-50!
dt[max_ct, on="mc", assigned_ct:=i.ct]
setkey(dt, assigned_ct, mc)
mc_annotation <- unique(dt[,.(mc,assigned_ct)])
setnames(mc_annotation, "assigned_ct", "ct")
```

Add temporary colours to metacells grouped by cell type.  

```{r mc_col}
require(RColorBrewer)

ct <- unique(dt$ct)
grep("i_neuron",ct, value = TRUE)
ct_neurons <- unlist(sapply(c("i_neuron_prog","i_neuron_ec","i_neuron_en"), grep, x=ct))
neur_palette <- colorRampPalette(brewer.pal(9, "Blues"))
neur_cols <- neur_palette(length(ct_neurons)+4)[c(F,F,rep(T,length(ct_neurons)),F,F)]
ct_others <- grep("i_neuron", ct, invert = TRUE)
others_palette <- colorRampPalette(brewer.pal(8, "Dark2"))
others_cols <- others_palette(length(ct_others))
mc_col <- data.table(
  ct = c(sort(ct[ct_neurons]), ct[ct_others]),
  col = c(neur_cols, others_cols)
)
mc_annotation[mc_col, on='ct', col:=i.col]
setkey(mc_annotation, ct)
```

Save cell and metacell annotations.  

```{r, annot_files_save}
fwrite(mc_annotation, mc_annotation_file, sep="\t")
fwrite(dt, mc_cell_info_file, sep="\t")
```

Barplot of metacells coloured by assigned cell type anottation.

```{r transfer_annot_plot}
require(ggplot2)
require(plotly)

mc_cols <- mc_annotation$col
names(mc_cols) <- mc_annotation$mc

niche_order <- mc_annotation$mc
dt[,mc:=factor(mc, levels=niche_order)]

ct_annotation <- unique(mc_annotation[,.(cell_type,color)])
ct_cols <- ct_annotation$col
names(ct_cols) <- ct_annotation$ct
ct_cols["unident"] <- "white"

gp <- ggplot(dt, aes(x=mc, fill=ct), colour="black") + 
  geom_bar(position = "fill") + 
  scale_x_discrete(breaks=dt$mc, labels=dt$mc) + 
  scale_fill_manual(values = ct_cols) +
  scale_x_discrete(labels = NULL) +
  theme_minimal() + theme(panel.grid = element_blank()) +
  labs(x="metacells",y="percent of cells in mc")
ggplotly(gp)
```


# Manual metacell solution 

Edit annotation file - reorder, reassign and kill some of the existing metacells.

```{r load_manual_annot}
mc_annotation_file <- file.path(scdb_dir, "hydra_metacell_annotation_manual.tsv")
mc_annotation <- fread(mc_annotation_file, sep="\t")
mc_cols <- mc_annotation$col
names(mc_cols) <- mc_annotation$mc
niche_order <- mc_annotation$mc
```

Update mc objects after manual reordering.  

```{r scdb_annot_update}
# remove outliers
out_mc <- 43
mc <- mc_set_outlier_mc(mc=mc, mc_ids=out_mc)
mc_filt <- mc_update_stats(mc=mc, scmat=mat)

# reorder metacells (mc_update_stats renames them to go from 1 to end, without gaps for removed mcs!)
niche_order_new <- niche_order
for (i in sort(out_mc)) {
  id_changed <- niche_order>i
  niche_order_new[id_changed] <- niche_order_new[id_changed]-1
}

mc_filt_ord <- mc_reorder(mc=mc_filt, ord=niche_order_new)

# add new mc object to db
# the mcs are now ordered and coloured by celltype and assigned names 1:ncol(mc_fp)
scdb_add_mc(id="mc_filt_ord", mc=mc_filt_ord)

# add filtered UMI matrix to db
mc_filt_ord_meta <- cbind(mc_filt_ord@mc)
rownames(mc_filt_ord_meta) <- names(mc_filt_ord@mc)
mat_filt_ord <- scm_new_matrix(
  mat=data.matrix(mat@mat[,names(mc_filt_ord@mc)]), 
  cell_metadata=mat@cell_metadata[names(mc_filt_ord@mc),], 
  stat_type="umi"
)
scdb_add_mat(id="mat_filt_ord", mat=mat_filt_ord)

# save annotation text files
mc_annotation_new <- data.table(
  metacell=names(mc@colors),
  color=mc@colors
)
setnames(mc_annotation,c("metacell","cell_type","color"))
mc_annotation_new[mc_annotation, on="color", "cell_type":=i.cell_type]
setcolorder(mc_annotation_new,c("metacell","cell_type","color"))
fwrite(mc_annotation_new, file.path(scdb_dir,"hydra_metacell_annotation_filt_ord.tsv"),sep="\t")

mc_cell_info <- data.table(
    cell=mc@cell_names,
    metacell=mc@mc
)
mc_cell_info <- mc_cell_info[mc_annotation, on="metacell"]
fwrite(mc_cell_info, file.path(scdb_dir, "hydra_cell_annotation_filt_ord.tsv"),sep="\t")
```

# Cell type annotations 

Load objects from metacell database

```{r transfer_annots_setup}
mc_id <- "mc_filt_ord"
mat_id <- "mat_filt_ord"
mc <- scdb_mc(mc_id)
mat <- scdb_mat(mat_id)
```

Create reordered and annotated metacell object.

```{r mc_reord_annot}
annk <- fread("metacell/Metacell_annotation_reord")
ctcol <- annk$color
names(ctcol) <- annk$metacell

# reorder
mc_reord <- mc_reorder(mc, as.integer(annk$metacell))
mc_reord@colors <- annk$color

# add mc object
if (all(!grepl("$Hvul_",rownames(mc_reord@mc_fp)))) {
  rownames(mc_reord@mc_fp) <- paste0("Hvul_",rownames(mc_reord@mc_fp))
  rownames(mc_reord@e_gc) <- paste0("Hvul_",rownames(mc_reord@e_gc))
}
scdb_add_mc("mc_090820",mc_reord)

# save final annots
ann_final <- file.path(scdb_dir,"Hvul_metacell_annotation")
fwrite(annk[,metacell:=1:.N], ann_final, sep="\t")
```

Remove biological doublets

```{r}
mc_id <- "mc_090820"
mat_id <- "mat_filt_ord"
mc <- scdb_mc(mc_id)
mat <- scdb_mat(mat_id)
ann <- fread("metacell/Hvul_metacell_annotation")
ann[grep("(pd|id|mp)$",cell_type),cell_type:="KILL"]
ann[,cell_type:=factor(cell_type,levels=c(setdiff(unique(ann$cell_type),"KILL"),"KILL"))]
new_ann <- ann[cell_type!="KILL"][,metacell:=1:.N]

mc_f=mc_set_outlier_mc(mc,as.integer(ann[cell_type=="KILL"]$metacell))
mc_f=mc_update_stats(mc_f,mat)
#scr_load_cell_type_table(file.path(scdb_dir,"Metacell_annotation_080420"),mc_f)
mc_f@colors=as.character(new_ann$color)
names(mc_f@colors)=as.character(new_ann$metacell)
scdb_add_mc("mc_100920",mc_f)
fwrite(new_ann,file.path(scdb_dir,"Hvul_metacell_annotation"),sep="\t",col.names=TRUE)
```

Save metacell output files.

```{r mc_save}
mc_id <- "mc_hydra"
mat_id <- "mat_filt_ord"
mc <- scdb_mc(mc_id)
mat <- scdb_mat(mat_id)
ann_file <- file.path(scdb_dir,"Hvul_metacell_annotation")
ann <- fread(ann_file)

write.table(
  mc@mc_fp, file=file.path(scdb_dir, "Hvul_metacell_gene_FC"), 
  sep="\t", quote=FALSE, row.names=TRUE, col.names=TRUE
)

mc_umis_list <- lapply(colnames(mc@mc_fp), function(x) {
  cells <- names(mc@mc)[mc@mc==x]
  Matrix::rowSums(mat@mat[,cells])
})
mc_umis <- do.call(cbind,mc_umis_list)
colnames(mc_umis) <- colnames(mc@mc_fp)
mc_umifrac <- mc_umis/rep(colSums(mc_umis),each=nrow(mc_umis))

write.table(
  mc_umis, file=file.path(scdb_dir, "Hvul_metacell_UMI_counts"), 
  sep="\t", quote=FALSE, row.names=TRUE, col.names=TRUE
)
write.table(
  mc_umifrac, file=file.path(scdb_dir, "Hvul_metacell_UMI_fractions"), 
  sep="\t", quote=FALSE, row.names=TRUE, col.names=TRUE
)

cell_df <- data.table(
  cell = mc@cell_names[!(mc@cell_names %in% mc@outliers)],
  metacell = mc@mc
)
fwrite(cell_df, file=file.path(scdb_dir, "Hvul_metacell_assignments"), sep="\t")

```

# Cell type level metacell analysis

Summarize metacells across cell types, and add to the database.

```{r ct_mc}
mc_id <- "mc_hydra"
mat_id <- "mat_filt_ord"
mat <- scdb_mat(mat_id)
mc <- scdb_mc(mc_id)
ann_file <- file.path(scdb_dir,"Hvul_metacell_annotation")
ann <- fread(ann_file)

# calculate mc object with mc_fp by cell type
mc_ct <- scr_cell_type_fp(
  input_table = ann_file,
  mc_object = mc,
  mat_object = mat
)
scdb_add_mc(id="ct_hydra",mc_ct)

# save footprint
write.table(
  mc_ct@mc_fp, file=file.path(scdb_dir, "Hvul_cell_type_gene_FC"), 
  sep="\t", quote=FALSE, row.names=TRUE, col.names=TRUE
)

# save mock ann
ann[,metacell:=cell_type]
# fwrite(unique(ann),file.path(scdb_dir,"Nvec_adult_cell_type_annotation"),sep="\t")

# counts
ct_umis_list <- lapply(colnames(mc_ct@mc_fp), function(y) {
  x <- ann[cell_type==y,]$metacell
  cells <- names(mc_ct@mc)[mc_ct@mc%in%x]
  Matrix::rowSums(mat@mat[,cells])
})
ct_umis <- do.call(cbind,ct_umis_list)
colnames(ct_umis) <- colnames(mc_ct@mc_fp)
ct_umifrac <- ct_umis/rep(colSums(ct_umis),each=nrow(ct_umis))

write.table(
  ct_umis, file=file.path(scdb_dir, "Hvul_cell_type_UMI_counts"), 
  sep="\t", quote=FALSE, row.names=TRUE, col.names=TRUE
)
write.table(
  ct_umifrac, file=file.path(scdb_dir, "Hvul_cell_type_UMI_fractions"), 
  sep="\t", quote=FALSE, row.names=TRUE, col.names=TRUE
)

cell_df <- data.table(
  cell = mc_ct@cell_names[!(mc_ct@cell_names %in% mc@outliers)],
  cell_type = mc_ct@mc
)
fwrite(cell_df, file=file.path(scdb_dir, "Hvul_cell_type_assignments"), sep="\t")
```

# Broad cell type level metacell analysis

Group  cell types into broad cell types, and add to the database.

```{r bct_mc}
#source("scr_plot_cmod_markers.R")
mc_id <- "ct_hydra"
mat_id <- "mat_filt_ord"
mat <- scdb_mat(mat_id)
mc <- scdb_mc(mc_id)
ann_file <- file.path(scdb_dir,"Hvul_broad_cell_type_annotation")
ann <- fread(ann_file)

# calculate mc object with mc_fp by cell type
mc_ct <- scr_cell_type_fp(
  input_table = ann_file,
  mc_object = mc,
  mat_object = mat
)
scdb_add_mc(id="bct_hydra",mc_ct)

# save footprint
write.table(
  mc_ct@mc_fp, file=file.path(scdb_dir, "Hvul_broad_cell_type_gene_FC"), 
  sep="\t", quote=FALSE, row.names=TRUE, col.names=TRUE
)

# counts
ct_umis_list <- lapply(colnames(mc_ct@mc_fp), function(y) {
  x <- ann[cell_type==y,]$metacell
  cells <- names(mc_ct@mc)[mc_ct@mc%in%x]
  Matrix::rowSums(mat@mat[,cells])
})
ct_umis <- do.call(cbind,ct_umis_list)
colnames(ct_umis) <- colnames(mc_ct@mc_fp)
ct_umifrac <- ct_umis/rep(colSums(ct_umis),each=nrow(ct_umis))

write.table(
  ct_umis, file=file.path(scdb_dir, "Hvul_broad_cell_type_UMI_counts"), 
  sep="\t", quote=FALSE, row.names=TRUE, col.names=TRUE
)
write.table(
  ct_umifrac, file=file.path(scdb_dir, "Hvul_broad_cell_type_UMI_fractions"), 
  sep="\t", quote=FALSE, row.names=TRUE, col.names=TRUE
)

cell_df <- data.table(
  cell = mc_ct@cell_names[!(mc_ct@cell_names %in% mc@outliers)],
  cell_type = mc_ct@mc
)
fwrite(cell_df, file=file.path(scdb_dir, "Hvul_broad_cell_type_assignments"), sep="\t")

```

# Sequencing and clustering statistics

Load objects from metacell database

```{r}
mc_id <- "mc_hydra"
mat_id <- "mat_filt_ord"
mat <- scdb_mat(mat_id)
mc <- scdb_mc(mc_id)

ann <- fread(file.path(scdb_dir,"Hvul_metacell_annotation"))
setnames(ann,c("metacell","cell_type","color"))
niche_order <- as.character(ann$metacell)
niche_color <- ann$color
```

Extract stats for metacell clustering.

```{r}
# cell sizes
cell_sizes <- apply(mat@mat,2,sum)
cdt <- data.table(cells=colnames(mat@mat),cell_sizes=cell_sizes)
cdt[data.table(cells=names(mc@mc),metacells=mc@mc), on="cells", mc:=i.metacells]
cdt <- cdt[!is.na(mc)]
cdt[,median_mc_cell_size:=median(cell_sizes),mc]
cdt[,mc:=factor(mc,levels=sort(as.integer(unique(cdt$mc))))]
cdt[,total_mc_umis:=sum(cell_sizes),mc]
cdt[,total_mc_cells:=.N,mc]
fwrite(cdt,file.path(scdb_dir,"cdt.tsv"),sep="\t")
# gene stats
gstat_id <- "gstat"
gstat <- scdb_gstat(gstat_id)
gset_id <- "clust_markers"
marker_genes_gset <- scdb_gset(gset_id)
markers <- names(marker_genes_gset@gene_set)
gstat <- setDT(gstat)
setnames(gstat,"name","gene")
gstat[, marker:="no"]
gstat[gene %in% markers, marker:="yes"]
gstat[,log2_ds_mean:=log2(ds_mean)]
fwrite(gstat,file.path(scdb_dir,"gstat.tsv"),sep="\t")
# markers in cells
markers_mat <- mat@mat[markers,cdt$cells]
markers_num <- apply(markers_mat, 2, function(x) sum(x>0))
mdf <- cbind.data.frame(num_markers=markers_num)
fwrite(mdf,file.path(scdb_dir,"mdf.tsv"),sep="\t")
```

Sequencing stats are from `Drop-seq_pipeline_initialQC/cell_estimation_initial_QC/Cell_QC_transcriptome.html` (on Dryad)

```{r}
sample_list <- c("D01-D1_S1","D01-P2_S4","D02-CO_S2", "D02-P1_S3", "D02-PB_S4", "D03-KI_S1", "D03-MA_S2", "D03-FM_S3","D06-FM_S1", "D06-MA_S3", "D06-KI_S4","D11-PO_S1","D11-BU_S2","D12-N1_S1","D12-N2_S2")

median_genes_str <- c("median genes 01-D1_ :3486", "median genes 01-P2_ :3122", "median genes 02-CO_ :1947", "median genes 02-P1_ :2382", "median genes 02-PB_ :2634", "median genes 03-KI_ :2474", "median genes 03-MA_ :2979", "median genes 03-FM_ :2747", "median genes 06-FM_ :3792", "median genes 06-MA_ :2104", "median genes 06-KI_ :2087", "median genes 11-PO_ :935", "median genes 11-BU_ :1342", "median genes 12-N1_ :509", "median genes 12-N2_ :477")
median_genes <- as.integer(str_extract(median_genes_str,"(?<=(\\:))\\d+"))

median_reads_str <- c("median reads 01-D1_ :37578","median reads 01-P2_ :24995","median reads 02-CO_ :13104","median reads 02-P1_ :18046","median reads 02-PB_ :17832","median reads 03-KI_ :33033","median reads 03-MA_ :35676","median reads 03-FM_ :41029","median reads 06-FM_ :34652","median reads 06-MA_ :17768","median reads 06-KI_ :18142","median reads 11-PO_ :22226","median reads 11-BU_ :20099","median reads 12-N1_ :29643","median reads 12-N2_ :19789")
median_reads <- as.integer(str_extract(median_reads_str,"(?<=(\\:))\\d+"))

median_umis_str <- c("median transcripts 01-D1_ :15873","median transcripts 01-P2_ :10618","median transcripts 02-CO_ :5242","median transcripts 02-P1_ :7048","median transcripts 02-PB_ :8206","median transcripts 03-KI_ :9990","median transcripts 03-MA_ :11762","median transcripts 03-FM_ :11091","median transcripts 06-FM_ :14300","median transcripts 06-MA_ :6372","median transcripts 06-KI_ :7349","median transcripts 11-PO_ :2261","median transcripts 11-BU_ :3261","median transcripts 12-N1_ :907","median transcripts 12-N2_ :895") 
median_umis <- as.integer(str_extract(median_umis_str,"(?<=(\\:))\\d+"))

cstat <- data.table(
  sample=sample_list, 
  mapped_reads_per_cell=median_reads, 
  genes_per_cell=median_genes, 
  UMI_per_cell=median_umis
)
fwrite(cstat, file.path(scdb_dir,"cellstat.tsv"), col.names = TRUE, sep = "\t")
```

Summarize

```{r}
cdt <- fread(file.path(scdb_dir,"cdt.tsv"))
cdt[,dataset:=str_extract(cells,"\\d{2}-[A-Z0-9]{2}")]
dt <- merge.data.table(
  cdt[,.(cells=.N),dataset],
  cdt[,.(UMI_per_cell=median(cell_sizes)),dataset],
  by="dataset"
)

counts <- fread(file.path(scdb_dir,"cellstat.tsv"))
counts[,dataset:=str_remove(str_remove(sample,"^D"),"_S\\d")][,sample:=NULL]

stats <- merge.data.table(dt,counts,by="dataset",all.x=TRUE)
stats[,mapped_reads_per_UMI:=median_reads/UMI_per_cell]
fwrite(stats,file.path(scdb_dir,"stats.tsv"),sep="\t")

```


# Session info

```{r session, eval=TRUE}
sessionInfo()
```

